Performance Comparison of Common Face Detection APIs using Video Stills from Professional Tennis Matches

### 2018-01-30 paper/paper.Rmd
rticles implementation, referencing and additional information

---
title: "Detecting Facial Expressions in Video Stills from Professional Tennis Matches"
author: "Stephanie Kobakian, Mitchell O'Hara-Wild, Dianne Cook, Stephanie Kovalchik"
date: "20 September 2017"
output: bookdown::pdf_document2
toc: false
fig_caption: yes
number_sections: false
fontsize: 11pt
documentclass: article
geometry: margin=1in
header-includes:
    - \usepackage{setspace}\doublespacing
bibliography: references.bib
biblio-style: chicago
link-citations: true
---

```{r cache=FALSE, echo=FALSE, message=FALSE, warning=FALSE, results='asis'}
#packages:
library(bookdown)
library(pander)
library(ggplot2)
library(knitcitations)
library(RefManageR)
library(readr)
library(knitr)
library(grid)
library(kfigr)
library(UpSetR)
library(descr)
library(xtable)
library(gtable)
library(ggthemes)
library(EBImage)
library(gridExtra)
library(ggmosaic)
library(tidyverse)
library(purrr)

knitr::opts_chunk$set(echo=FALSE, message=FALSE, warning=FALSE, kfigr.link=TRUE, kfigr.prefix=TRUE, cache=TRUE, fig.env = TRUE, fig.cap=TRUE, fig.height=3)

options("citation_format" = "pandoc")

BibOptions(check.entries = FALSE, style = "markdown", bib.style = "alphabetic", cite.style = 'alphabetic')

#csv files:
amin<-read_csv("data/ALLmetaIMGnamed.csv", col_types = cols(type = col_factor(levels = c("Manual", "Animetrics", "Google", "Microsoft", "Skybiometry")))) %>% 
  filter(!duplicates)
SceneAttributes<-read_csv("data/SceneAttributes.csv")
FaceAttributes<-read_csv("data/FaceAttributes.csv")

@@

The report outlined that Animetrics [@Animetrics] required an 'image/face proportion should be at least 1:8 and that at least 64 pixels between eyes are required'. 
SkyBiometry [@Skybiometry] is a 'spin-off of Neurotechnology' which was considered by Gorodnichy et al. [-@Survey].
Companies who have recently expanded their API ranges were also considered. This search produced two of the software, Microsoft API [@Microsoft], provided by Microsoft Cognitive Services, and Google Vision API [@Google].
Online demos were used to test viability, Figure \@ref(fig:Trial-Image) depicts Bernard Tomic and the detected areas found.

@@

```{r Scene-Attributes-Table, fig.cap ="\\label{tab:Scene-Attributes-Table}"}
knitr::kable(SceneAttributes, format = "latex", longtable=FALSE, booktabs =TRUE, caption = "Characteristics of the scene occuring when the still shot was taken, one option for each category was selected for each image.")
```

@@

Table \@ref(tab:Face-Attribute-Tables) summarises attribute information recorded for each face.

```{r Face-Attribute-Tables, results='asis', fig.cap = "\\label{tab:Face-Attribute-Tables}"}
knitr::kable(FaceAttributes, format = "latex", longtable=FALSE, booktabs = TRUE, caption = "Characteristics of the faces captured, one option for each category was selected for each face manually tagged.")
```

@@

Figure \@ref(fig:timeFigure) shows the distribution of processing time required by the APIs for each image. The violin plot display created by @Violin 'combines the box plot and the density'. A categorical variable splits the data along the x axis, as would be seen in a box plot display. A density plot is then incorporated on the y axis. This is produced by manipulating a density display; switching the axes allows the continuous variable to increase along on the y axis, then the violin shape is created by mirroring the density. The width of the violin will now be twice the length that would be have been measured vertically in a density plot.

As seen in the facets of Figure \@ref(fig:timeFigure), both Microsoft and Google had few images which took a significant amount of time to complete, and these were removed from this plot due to concerns regarding the network. The times taken to search are quite variable between APIs, with Skybiometry consistently spending less time than other APIs. Microsoft and Google's times vary regardless of the number of faces found. The lines overlaid in each plot represent the smoothed linear model of the amount of faces, it can be seen that all APIs take longer to complete detection of faces in an image when there are more faces. As all APIs had more images where only one or two faces were found it resulted in more variability in the amount of time taken.

@@

```{r timeFigure, fig.height=6, fig.cap="System time taken by the four APIs to finish searching individual images, by number of faces discovered, displayed as violins. The width of the violin shows the most common times for the group of images.\\label{fig:timeFigure}"}

@@

knitr::kable(timetable, booktabs = TRUE, longtable=TRUE, caption="This table allows comparisons to be made between the  amount of time taken for each API took to detect faces in each image. Most of the images take less than one second to process. Google had the highest maximum time of 4.936.", format = "latex") #%>%
 # add_header_above(c(" " = 1, "Group 1" = 5, "Group 2" = 5))
```

@@

knitr::kable(ModelEstimatesTable, format = "latex", align=c('l', 
                                           'r', 'l',
                                           'r', 'l',
                                           'r', 'l',
                                           'r', 'l'),
             booktabs = TRUE, longtable=FALSE, caption="Evaluation of factors affecting face recognition for API, as assessed by logistic regression model. Coefficients from the four models shown (significance: ** 0.01 * 0.05 . 0.01 - NS). Most APIs wre affected by similar factors, with glasses, visor or hat, head angle, lighting, shot angle and situation all contributing significantly to face detection.")
```

@@

Table \@ref(tab:facecounts) displays the proportion of the manually tagged faces that the APIs detected. Google had by far the best detection rate. In addition, the APIs also detected non-faces, in images that produced no manually tagged faces. Google had the highest of these, and the detections were for the most part helpful, primarily of tiny faces in the crowd, that were ignored by manual tagging. They would not be needed for the long term purpose of the study, to detect and tag players' emotions during a match.

@@

Figure \@ref(fig:FaceUpset) is a set visualization to examine the intersection of the APIs, and the manual annotations. This is produced with the UpSetR [@UpSetR] package. The black bubbles below the bars indicate the API combination that corresponds to the bar count. The largest count comes from manual tagging, followed by Google and manual tagging, and then Google. Faces detected by all four API and manual annotations were the fourth largest count. This says that the APIs agreed on about a quarter of the faces from the manually tagged collection. The fifth group was the most surprising -- these were faces captured by Animetrics alone, that were false discoveries.

@@

Animetrics provided many surprises. Figure \@ref(fig:UnusualImages) shows four images and the captures where non-faces were detected as faces. In image (A) the player's back is detected as a face.
In image (B) the KIA logo was reported as a face. 
In image (C), where logos on the player's and ball boy's shirts were returned as faces. In image (D) where there are a lot of crowd faces, if you look closely one of the detections is actually a fist (to the right on the center).

@@

```{r Image-Characteristics-Table, fig.cap ="\\label{tab:Image-Characteristics-Table}"}
ImageCharacteristics <- amin %>% filter(type=="Manual") %>% group_by(situation, bg, shotangle, detect) %>% dplyr::summarise(count=n())
ImageCharacteristics<-arrange(ImageCharacteristics, desc(count))
ImageCharacteristics<-ImageCharacteristics[1:5,]
knitr::kable(ImageCharacteristics, booktabs = TRUE, longtable=FALSE, format.args=list(width=8), caption="Top five most frequent image attribute combinations.", format = "latex")
```

```{r ICFigs, fig.width = 8, fig.height = 8, fig.cap = "Relationships between court and face characteristics, and API detection, using mosaic plots: (A) Angle of face capture, (B) court situation, (C) angle of capture by background, and (D) situtation of capture by background. Most face captures occurred at a level angle, and this produces slightly better detection. Most situtation face captures are close up, but the detection rate is identical regardless of situation. Most level angles had the logo wall in the background, while most birds-eye had the court as background, as we would expect. Close ups were mostly against the logo wall. \\label{fig:ICFigs}"}

@@

Image characteristics (Table \@ref(tab:Scene-Attributes-Table)) were constructed during manual tagging. The scene information was tallied and the most common combinations of these features shown in Table \@ref(tab:Image-Characteristics-Table).

Figure \@ref(fig:ICFigs) examines the relationships between several of the image characteristics, and face detection by the Google API, using mosaic plots. Plot (A) compares the angle of capture by detection rate. Color is mapped to detection. Most images were taken at a level angle. The highest capture rate occurred when the angle was level, and least with a bird's eye, as might be expected. Plot (B) compares the situation with face detection. Close-up was the most common situation. The detection rate was the same across situations.

The mosaic \@ref(fig:ICFigs) (B) in Figure \@ref(fig:ICFigs), shows the amount of Faces captured during each possible situation. It contrasts the proportion of the images captured in each situation that either did or did not match the faces annotated manually. It can be seen that the largest amount of the faces were captured in a close up situation and there were more of these faces that were found by the APIs that did match manually derived faces. This proportion is steady across most situations, it can be inferred the situation may not be influencing the APIs rate of detection.

@@

The use of accessories like glasses and head wear, visors or hats, is common during the Australian Open as play is mostly outdoors during daytime.  Table \@ref(tab:accessories) shows player use of head wear and glasses, as tagged manually. Most players do not have glasses, but head wear, either visors or hats, are very common.

@@

Table \@ref(tab:glassesheadwear) compares the face capture rate of the different APIs for the player accessories, as compared to the manually tagged face attributes. Conditional proportions are calculated two ways: (1) proportion of the manually tagged faces, (2) proportion detected with the accessory relative to without. From the former conditional probability it can be seen that Google detects more faces regardless of the accessory, with slightly higher proportion when the player has no head wear or glasses. From the latter calculation, Google has less difference between the accessory or not detection rate. Overall, Google detects a higher proportion of faces, but they also have better detection rates if faces of players wearing head wear and glasses. Animetrics performed the worst when a player was wearing head wear or glasses.

>>>>

---
title: "Performance Comparison of Common Face Detection APIs using Video Stills from Professional Tennis Matches"

# to produce blinded version set to 1
blinded: 0

authors: 
- name: Stephanie Kobakian
  thanks: The authors gratefully acknowledge Tennis Australia.
  affiliation: Monash University

- name: Mitchell O'Hara-Wild
  affiliation: Monash University

- name: Dianne Cook
  affiliation: Monash University

- name: Stephanie Kovalchik
  affiliation: Victoria University and Tennis Australia

keywords:
- data science
- sports analytics
- shiny
- R

abstract: |
  This paper examines the effectiveness of facial detection APIs on broadcast video stills of Australian Open tennis matches. The goal is to determine the best API to use for face detection of players throughout a match. For training purposes, faces were manually tagged in 6406 images, using a specially constructed web application, recording the scene characteristics and features of individual faces, such as head wear or sunglasses. API performance was evaluated on their success rate at detecting the tagged faces, and also relative face and scene attributes.

toc: false
fig_caption: yes
number_sections: false
fontsize: 11pt
header-includes:
    - \usepackage{booktabs}
    - \usepackage{longtable}
bibliography: references.bib
biblio-style: chicago
output: rticles::asa_article
link-citations: true
---

```{r cache=FALSE, echo=FALSE, message=FALSE, warning=FALSE, results='asis'}
#packages:
library(bookdown)
library(pander)
library(ggplot2)
library(knitcitations)
library(RefManageR)
library(readr)
library(knitr)
library(grid)
library(kfigr)
library(UpSetR)
library(descr)
library(xtable)
library(gtable)
library(ggthemes)
library(EBImage)
library(gridExtra)
library(ggmosaic)
library(tidyverse)
library(purrr)
library(rticles)

knitr::opts_chunk$set(echo=FALSE, message=FALSE, warning=FALSE, kfigr.link=TRUE, kfigr.prefix=TRUE, cache=TRUE, fig.env = TRUE, fig.cap=TRUE)

options("citation_format" = "pandoc",knitr.table.format = "latex")

BibOptions(check.entries = FALSE, style = "markdown", bib.style = "alphabetic", cite.style = 'alphabetic', longnamesfirst=T)

#csv files:
amin <-read_csv("data/ALLmetaIMGnamed.csv", col_types = cols(type = col_factor(levels = c("Manual", "Animetrics", "Google", "Microsoft", "Skybiometry")))) %>% 
  filter(!duplicates)
SceneAttributes<-read_csv("data/SceneAttributes.csv")
FaceAttributes<-read_csv("data/FaceAttributes.csv")

@@

The report outlined that Animetrics [@Animetrics] required an 'image/face proportion should be at least 1:8 and that at least 64 pixels between eyes are required'. 
SkyBiometry [@Skybiometry] is a 'spin-off of Neurotechnology' which was considered by Gorodnichy et al. [-@Survey].
Companies who have recently expanded their API ranges were also considered. This search produced two of the software, Microsoft API [@Microsoft], provided by Microsoft Cognitive Services, and Google Vision API [@Google].
Online demos were used to test viability, Figure \ref{fig:Trial-Image} depicts Bernard Tomic and the detected areas found.

@@

```{r Scene-Attributes-Table, fig.cap ="\\label{tab:Scene-Attributes-Table}"}
kable(SceneAttributes, format = "latex", longtable=FALSE, booktabs =TRUE, caption = "Characteristics of the scene occuring when the still shot was taken, one option for each category was selected for each image.")
```

@@

Table \ref{tab:Face-Attribute-Tables} summarises attribute information recorded for each face.

```{r Face-Attribute-Tables, results='asis', fig.cap = "\\label{tab:Face-Attribute-Tables}"}
kable(FaceAttributes, format = "latex", longtable=FALSE, booktabs = TRUE, caption = "Characteristics of the faces captured, one option for each category was selected for each face manually tagged.")
```

@@

Figure \ref{fig:timeFigure} shows the distribution of processing time required by the APIs for each image. The violin plot display created by @Violin 'combines the box plot and the density'. A categorical variable splits the data along the x axis, as would be seen in a box plot display. A density plot is then incorporated on the y axis. This is produced by manipulating a density display; switching the axes allows the continuous variable to increase along on the y axis, then the violin shape is created by mirroring the density. The width of the violin will now be twice the length that would be have been measured vertically in a density plot.

As seen in the facets of Figure \ref{fig:timeFigure}, both Microsoft and Google had few images which took a significant amount of time to complete, and these were removed from this plot due to concerns regarding the network. The times taken to search are quite variable between APIs, with Skybiometry consistently spending less time than other APIs. Microsoft and Google's times vary regardless of the number of faces found. The lines overlaid in each plot represent the smoothed linear model of the amount of faces, it can be seen that all APIs take longer to complete detection of faces in an image when there are more faces. As all APIs had more images where only one or two faces were found it resulted in more variability in the amount of time taken.

@@

```{r timeFigure, fig.cap="System time taken by the four APIs to finish searching individual images, by number of faces discovered, displayed as violins. The width of the violin shows the most common times for the group of images.\\label{fig:timeFigure}"}

@@

kable(timetable, booktabs = TRUE, longtable=TRUE, caption="This table allows comparisons to be made between the  amount of time taken for each API took to detect faces in each image. Most of the images take less than one second to process. Google had the highest maximum time of 4.936.", format = "latex") #%>%
 # add_header_above(c(" " = 1, "Group 1" = 5, "Group 2" = 5))
```

@@

kable(ModelEstimatesTable, format = "latex", align=c('l', 
                                           'r', 'l',
                                           'r', 'l',
                                           'r', 'l',
                                           'r', 'l'),
             booktabs = TRUE, longtable=FALSE, caption="Evaluation of factors affecting face recognition for API, as assessed by logistic regression model. Coefficients from the four models shown (significance: ** 0.01 * 0.05 . 0.01 - NS). Most APIs wre affected by similar factors, with glasses, visor or hat, head angle, lighting, shot angle and situation all contributing significantly to face detection.")
```

@@

Table \ref{tab:facecounts} displays the proportion of the manually tagged faces that the APIs detected. Google had by far the best detection rate. In addition, the APIs also detected non-faces, in images that produced no manually tagged faces. Google had the highest of these, and the detections were for the most part helpful, primarily of tiny faces in the crowd, that were ignored by manual tagging. They would not be needed for the long term purpose of the study, to detect and tag players' emotions during a match.

@@

Figure \ref{fig:FaceUpset} is a set visualization to examine the intersection of the APIs, and the manual annotations. This is produced with the UpSetR [@UpSetR] package. The black bubbles below the bars indicate the API combination that corresponds to the bar count. The largest count comes from manual tagging, followed by Google and manual tagging, and then Google. Faces detected by all four API and manual annotations were the fourth largest count. This says that the APIs agreed on about a quarter of the faces from the manually tagged collection. The fifth group was the most surprising -- these were faces captured by Animetrics alone, that were false discoveries.

@@

Animetrics provided many surprises. Figure \ref{fig:UnusualImages} shows four images and the captures where non-faces were detected as faces. In image (A) the player's back is detected as a face.
In image (B) the KIA logo was reported as a face. 
In image (C), where logos on the player's and ball boy's shirts were returned as faces. In image (D) where there are a lot of crowd faces, if you look closely one of the detections is actually a fist (to the right on the center).

@@

```{r Image-Characteristics-Table, fig.cap ="\\label{tab:Image-Characteristics-Table}"}
ImageCharacteristics <- amin %>% filter(type=="Manual") %>% group_by(situation, bg, shotangle, detect) %>% dplyr::summarise(count=n())
ImageCharacteristics<-arrange(ImageCharacteristics, desc(count))
ImageCharacteristics<-ImageCharacteristics[1:5,]
kable(ImageCharacteristics, booktabs = TRUE, longtable=FALSE, format.args=list(width=8), caption="Top five most frequent image attribute combinations.", format = "latex")
```

```{r ICFigs, fig.cap = "Relationships between court and face characteristics, and API detection, using mosaic plots: (A) Angle of face capture, (B) court situation, (C) angle of capture by background, and (D) situtation of capture by background. Most face captures occurred at a level angle, and this produces slightly better detection. Most situtation face captures are close up, but the detection rate is identical regardless of situation. Most level angles had the logo wall in the background, while most birds-eye had the court as background, as we would expect. Close ups were mostly against the logo wall. \\label{fig:ICFigs}"}

@@

Image characteristics (Table \ref{tab:Scene-Attributes-Table}) were constructed during manual tagging. The scene information was tallied and the most common combinations of these features shown in Table \ref{tab:Image-Characteristics-Table}.

Figure \ref{fig:ICFigs} examines the relationships between several of the image characteristics, and face detection by the Google API, using mosaic plots. Plot (A) compares the angle of capture by detection rate. Color is mapped to detection. Most images were taken at a level angle. The highest capture rate occurred when the angle was level, and least with a bird's eye, as might be expected. Plot (B) compares the situation with face detection. Close-up was the most common situation. The detection rate was the same across situations.

The mosaic \ref{fig:ICFigs} (B) in Figure \ref{fig:ICFigs}, shows the amount of Faces captured during each possible situation. It contrasts the proportion of the images captured in each situation that either did or did not match the faces annotated manually. It can be seen that the largest amount of the faces were captured in a close up situation and there were more of these faces that were found by the APIs that did match manually derived faces. This proportion is steady across most situations, it can be inferred the situation may not be influencing the APIs rate of detection.

@@

The use of accessories like glasses and head wear, visors or hats, is common during the Australian Open as play is mostly outdoors during daytime.  Table \ref{tab:accessories} shows player use of head wear and glasses, as tagged manually. Most players do not have glasses, but head wear, either visors or hats, are very common.

@@

Table \ref{tab:glassesheadwear} compares the face capture rate of the different APIs for the player accessories, as compared to the manually tagged face attributes. Conditional proportions are calculated two ways: (1) proportion of the manually tagged faces, (2) proportion detected with the accessory relative to without. From the former conditional probability it can be seen that Google detects more faces regardless of the accessory, with slightly higher proportion when the player has no head wear or glasses. From the latter calculation, Google has less difference between the accessory or not detection rate. Overall, Google detects a higher proportion of faces, but they also have better detection rates if faces of players wearing head wear and glasses. Animetrics performed the worst when a player was wearing head wear or glasses.
